The markings of the trade: exploring the patterns of olive oil production in Roman Baetica

### 2020-07-23 Paper/jaccardbaetica.tex
add new changes xavi

Some hypotheses argue elsewhere that this same \textit{Annona} could also have organized the supply of additional goods such as olive oil to the Roman legions \citep{remesal_annona_1986,
remesal_annona_1990}

@@

An explanation for the interest to export olive oil for the military garrison in the Roman provinces is still unclear. The origin of this exportation may be related to two aspects: a) cultural consumption whose olive oil is used by cultural reasons such as group identification or by habit and b) economical and political reasons to create a redistributive system mechanism to supply military camps.

@@

Julius Caesar's campaigns in \textit{Gallia} during 1st BC opened the door to the invasion of \textit{Germania} during Augustus rule~\citep{remesal_annona_1986,remesal_baetica_2002}\xavi{no hay un resumen de la conquista similar para Britannia, queda raro}. Previous studies suggested that the supply to the German limes was mainly based on riverine transport, but recent works suggest that the maritime route through the Atlantic Ocean could have been more important than expected (both for Germania and Britannia).~\citep{remesal_germn_2010,rubio-campillo_ecology_2018}.\xavi{lo del océano atlántico está un poco descontextualizado; no sería mejor ponerlo al final de esta sección?}

Roman studies exploring the presence of Dressel 20 amphorae in the provinces Inferior and Superior have not had the same impact as the rest of the provinces.
%\xavi{no entiendo esta frase. Que qué provincia? Germania son 2 provincias no?}

This could be explained due to the lack of archaeological sources \citep{horacio2010llegada}. As a consequence, it is still unknown if trade agents participated on the distribution of Baetican olive oil in \textit{Germania} or the supply was exclusively focused on the Roman army garrisons\citep[156]{remesal_germn_2010}\xavi{quieres decir que aquí sólo encontramos dressel 20 en yacimientos militares porque no se han excavado muchos civiles?}.

@@

The presence of the Roman army encouraged the exchange in the province as shown by the arrival of this product both civil settlement and military sites with a mayor concentration at the German \textit{limes}\xavi{pero justo antes decías que no había en civiles no?}. It seems that some Baetican centres would be assigned to the support of olive oil\xavi{the support of olive oil no se entiende. Además, falta alguna cita y desarrollar esta idea}. However, this hypothesis is hard to assess given the current lack of archaeological evidence related to 
\xavi{related to BLABLABLA. Habría que detallar un poco porque hay un montón de evidencia arqueológica del limes, así que qué es exactamente lo que falta?}\citep[125]{remesal_concierto}.

>>>>

However, some hypotheses argue elsewhere that this same \textit{praefectura annonae} could also have organized the supply of additional goods such as olive oil to the Roman legions  \citep{remesal_annona_1986,
remesal_annona_1990}

@@

An explanation for the interest to export olive oil for the military garrison in the Roman provinces is still unclear. The origin of this exportation may be related to two aspects: a) cultural consumption whose olive oil is used by cultural reasons such as group identification or by habit and b) economical and political reasons to create a redistributive system mechanism to supply military camps. 
\maria{esta frase está explicada mejor}

@@

%Julius Caesar's campaigns in \textit{Gallia} during 1st BC opened the door to the invasion of \textit{Germania} during Augustus rule~\citep{remesal_annona_1986,remesal_baetica_2002}\xavi{no hay un resumen de la conquista similar para Britannia, queda raro}. 
\maria{lo he eliminado}

Roman studies exploring the presence of Dressel 20 amphorae in the provinces Inferior and Superior during Roman Empire have not had the same impact as the rest of the provinces \citep[293]{remesal_baetica_2002}.
%\xavi{no entiendo esta frase. Que qué provincia? Germania son 2 provincias no?}

The massive importation of Dressel 20 in both \textit{Germania} provinces could be explained due to the archaeological sources found mostly in German \textit{limes}. 
By contrast, it is still unknown if trade agents participated on the distribution of Baetican olive oil in \textit{Germania} or the supply was exclusively focused on the Roman army garrisons\citep[156]{remesal_germn_2010}
\xavi{quieres decir que aquí sólo encontramos dressel 20 en yacimientos militares porque no se han excavado muchos civiles?}.
\maria{lo he mejorado a ver si se entiende}

%This could be explained due to the lack of archaeological sources \citep{horacio2010llegada}

@@

The presence of the Roman army encouraged the exchange in the province as shown by the arrival of this product both civil settlement and military sites with a mayor concentration at the German \textit{limes}.

\xavi{pero justo antes decías que no había en civiles no?}. 
\maria{yo no he dicho eso :-[}

It seems that some Baetican centres could have been assigned to a specific province for the olive oil supply to military garrisons \citep[125]{remesal_concierto}. 
\xavi{the support of olive oil no se entiende. Además, falta alguna cita y desarrollar esta idea}. 
\maria{mejorado, creo}
However, this hypothesis is hard to assess given the current lack of written evidence related to the meaning of the amphoric stamps. 
\xavi{related to BLABLABLA. Habría que detallar un poco porque hay un montón de evidencia arqueológica del limes, así que qué es exactamente lo que falta?}. 
\maria{mejorado,creo}

Previous studies suggested that the supply to the German limes was mainly based on riverine transport, but recent works suggest that the maritime route through the Atlantic Ocean could have been more important than expected (both for Germania and Britannia).~\citep{remesal_germn_2010,rubio-campillo_ecology_2018}.
%\xavi{lo del océano atlántico está un poco descontextualizado; no sería mejor ponerlo al final de esta sección?}
